---
title: "esm244_video_lab4_LB"
author: "Lydia Bleifuss"
date: "1/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#Header, put a space after the # to make header font, with more #, smaller headers, any text with out a # in front is normal text

### 1. Attach necessary packages

We'll be using the **tidyverse**, **sf**, **tmap** packages! 

** on either side of text makes it bold!

Command + Option + I to add a code chunk

```{r, include = FALSE}
## above, to make code messaging not show up, in {r, include = FALSE}, now nothign from this code chunk shows up, messages = FALSE, and echo = FALSE area also helpful (also can look at Help, Rmarkdown cheat sheet to see these different options)

library(tidyverse)
library(sf)
library(tmap)
```

### 2. Read in CA eco-regions data

```{r, message = FALSE}
#we want to read all of the parts of ca_eco all at once, all contain parts of spatial data, need read_sf package!

ca_eco <- read_sf(dsn = "session_3_materials", layer = "ca_eco") %>% #geometry column looks like a disaster but contains all the needed spatial data (polygons, points, etc) and it is sticky (it holds onto/is retained no matter what cleaning/organizing you are doing)
  select(US_L3NAME) %>% 
  rename(region = US_L3NAME) %>% #going to simplify here, decreasing number of nodes that make up the outline of each polygon (probablyl would not do this in reality, will run faster in the lab)
  st_simplify(dTolerance = 100) %>% 
  st_transform(crs = 4326)

plot(ca_eco) #just basic viz but already cool! 


#please note, in the rmapshaper::ms_simplify(), and geos::gSimplify also reshape polygons, there are options 
```

**You can always check the projections/coordinate reference system (CRS) (in console, use st_crs(sf object name) or in this case st_crs(ca_eco), we can set the projection by piping onto ca_eco!** 

### 3. Read in CA Counties polygons

```{r}

ca_counties <- read_sf(dsn = "session_3_materials", layer = "california_county_shape_file")

st_crs(ca_counties) = 4326 #since there wasn't an existing crs, set it to be 4326


```

### 4. Read in CA dams locations & data

```{r}

ca_dams <- read_sf(dsn = "session_3_materials", layer = "California_Jurisdictional_Dams")

#in geometry, just has lat and long because it is point data

#checked crs in console and is already set to 4326! Good to go 
```

### 5. Make a map of dams in California, overlayed on eco-regions, with outlines of CA counties 

```{r}
#ggplot works well with spacial data (needs to know ggplot, data, and geometry to produce)

#don't give x and y with spatial data because it recognizes lat and long from geometry column 

ggplot(data = ca_counties) +
  geom_sf(color = "black",
          size = 0.1) + #you are only changing things in aes (aestetics) when they depend on a a specific variable in the dataset 
  geom_sf(data = ca_eco,
          aes(fill = region), #lower the alpha gets from 1, the more transparent it will be, color is for the outlines and with NA there is no outline
          alpha = 0.5, color = NA,
          show.legend = FALSE) +
  geom_sf(data = ca_dams,
          size = 0.5,
          alpha = 0.5) +
  theme_minimal()
```

### 6. Just explore SB county eco-regions

```{r}
#going to look at SB county, so where Name = Santa Barbara  (use filter to keep based on rows)

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara")

# Clip eco-region data to only include information within SB county - use st_intersection() 

eco_clip <- st_intersection(ca_eco, sb)

# Now let's plot it! Good to check in on what happened: ggplot2

ggplot() +
  geom_sf(data = ca_counties, 
          fill = "gray90", 
          color = "gray80",
          size = 0.2) +
  geom_sf(data = eco_clip,
          aes(fill = region),
          color = "white",
          size = 0.4,
          show.legend = FALSE) +
  coord_sf(xlim = c(-121, -119), ylim = c(33.5, 35.5)) + #can change a lot with coord_sf 
  scale_fill_manual(values = c("magenta", "gold2", "blue")) 
```


### 7. Create an interactive map with tmap

```{r}
# go to leaflet-extras.github.io (BASE MAPS!!!)

sb_clip_tmap <- tm_basemap("Esri.WorldImagery") +
  tm_shape(eco_clip) + 
  tm_fill("region", palette = c("orange", "purple", "yellow"), alpha = 0.5)

#to make interactive need to net to tmap_mode("view"")

tmap_mode("view")
sb_clip_tmap

```

### 8. How to create an 'sf' object from lat/long recording (e.g. in an Excel file) THIS IS IMPORTANT!! 

Make a mock dataset using tribble():

```{r}

my_example <- tribble(
  ~id, ~lon, ~lat, #these are the three columns
  "tiger", -119.4, 34.35,#these are the rows full o of data 
  "lion", -119.41, 34.39,
  "bear", -119.43, 34.38
)

#just because you create columns that say long and lat, does not mean that r knows it's spatial, you have to tell it! 

# Convert this to 'sf' object:

animals_sf <- st_as_sf(my_example, coords = c("lon", "lat"), crs = 4326) #now, r has created the geometry column with the lat and long coords! 

```

Make a map using tmap!

```{r}

animal_map <- tm_basemap("Stamen.Watercolor") +
  tm_shape(animals_sf) +
  tm_dots(labels = "id", col = "purple", size = 0.5) #using tm_dots because they are points

#tmap is already in view mode based on above code chunk

animal_map

```


### 8. Cloropleth of dam counts

```{r}

intersection <- st_intersection(x = ca_dams, y = ca_counties)

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()

#want to merge count data with county geometry data 
ca_tot <- ca_counties %>% 
  st_join(dams_per_county) %>% #we now see a NAME.x, which just means that it didn't want to copy over teh othr name column because they were slightly different
  select(NAME.x, n) %>% 
  rename(name = NAME.x)

ca_tot_zero <- ca_tot %>% 
  replace_na(list(n = 0))

# One mor map example with ggplot:

ggplot() +
  geom_sf(data = ca_tot_zero,
          aes(fill = n),
          size = 0.2,
          color = "white") +
  scale_fill_continuous(low = "yellow", high = "red")

```

Vignettes for sf (google sf package in Github) - great resource!!

Fee book: "Geocomputation with R" by Robin Lovelace, super helpful!

